Speculation in foreign exchange: noise or information

### 2015-06-15 Knitr/NewPaper.Rnw
Updated most of the paper. The final couple of sections need to be completed. The whole paper could be made more dramatic with the key event charts mroe towards the front.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself as the evolution of collective sentiment encourages trends in asset prices.

Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.  The cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration. The Kyle model has market-makers reciving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange \citet{Kyle1985Continuous} and \citet{glosten1985bid}.  Here the speculators may be informed or uninformed.

% does this following paragraph come later?  I think so....Once the three
% main categories are done. 
In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is essential to encourage te process of price discovery. 
%
Fisher Black suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

  Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.  \citet{muth1961Rational} uses commodity speculation to argue for model-consistent expectation formation.  In this case, the speculators are informed and tend to reduce the volatility of prices.

@@

\subsection*{The problem of uninformed noise-traders}
The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Other empirical work}
There is little empirical eassessment of these theories about speculation and noise-trading. In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.

@@

\subsection{Why this is important}
The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial by reducing risk and encouraging more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   This will involve three steps:  first,assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

The noise-trader model used in this study is from De Long, Shleifer, Summers and Waldeman \citep{Delong1990noise}. This is a derivative of the Samuelson overlapping generations model  with two-period agents making a portfolio choice \citep{Samuelson1958}.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r)^2}
\label{eq:NP2}
\end{equation}

The following effects are noted:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found in day-traders \citet{odean1998volume}

\subsection{The implications of the noise-trader model}
The aim of this study is to test the implications of the noise-trader model.  There are two parameters from the noise-trader price Equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

@@

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The package \emph{eventstudies} \citep{eventstudies} aligns the defined events (in this case the extreme noise-trader activity) so that descriptive statistics and other analysis can be carried out on the period after the event has occurred.  Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\begin{table}[h!]
\begin{threeparttable}
\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
\begin{tabular}{p{6cm} | r r r r}
 \label{tabref:MC1}
                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
  \hline
\hline
Mean                     & -0.00 & -0.90*   & -0.49 \\
Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
Skew                      &  -0.07 &  -0.16   &  -0.78*\\
Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
Number                 &  1000        &   50        & 50\\

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}    This is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

Determining the fundamental value of exchange rates has proved to be extremely difficult.  As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.  Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception. \footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account. Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}

@@

% The table may be pushed onwards.  Use t (i think). 
\begin{landscape}
\begin{table}[t]
\begin{threeparttable}
\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
  \hline
Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
\hline
 \label{tabref:DS1}
\end{tabular}
\begin{tablenotes}
\small
\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY). The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.    
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. This apparent stability hides some significant changes that are taking place below the surface.  For example, the Japanese yen records an appreciation of 10.66\% on one day.  Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.  An asterisk is used to denote statistical significance at the level of 95\% confidence.  A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.

@@

The CFTC require traders to categorise themselves as being either commercial (C) or non-commercial (NC).  Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.  These could be seen as natural hedgers.  Non-commercial accounts are generally considered to be speculators.  See \citep{FuturesSanders} and \citep{FuturesWang} for examples of previous use of this data in this fashion.

@@

It could be argued that this is a biased sample as it is a US market.  Though international financial markets are global in nature, there is some evidence to support this claim.  For example, there is a tendency to have a positive holding of the overseas currency as this is a hedge for exporters.  Nonetheless, as was argued with the risk-reversals, it is the relative rather than absolute position of the signal that will determine the extreme and therefore if the US sample is proportionate to the population, sample signals can still be regarded as measuring an extreme.  If the distribution of the sample itself is different from the population this is a problem, but there is no evidence that this is the case.

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself as the evolution of collective sentiment encourages trends in asset prices.

Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.  The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration. The \citet{Kyle1985Continuous} model has market-makers reciving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange.  See also \citet{glosten1985bid}.

% does this following paragraph come later?  I think so....Once the three
% main categories are done. 
In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

  Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.  \citet{muth1961Rational} uses commodity speculation as the means to drive prices towards the \emph{rationa-expectations-equilibrium} (RER).  In this case, the speculators are informed and tend to reduce the volatility of prices. This is a reiteration of the Friedman idea of speuclation but runs immediately into the \citet{milgrom1982information} problem of finding counterparties willing to make the other side of the transaction.

@@

%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Other empirical work}
There is little empirical eassessment of these theories about speculation and noise-trading. In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.

@@

\subsection{Why this is important}
The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial by reducing risk and encouraging more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   This will involve three steps:  first, assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

The \citep{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citep{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}
\label{eq:NP2}
\end{equation}

The following effects are noted:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.

\subsection{The implications of the noise-trader model}
There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  the aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

@@

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\begin{table}[h!]
\begin{threeparttable}
\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
\begin{tabular}{p{4cm} | r r r r}
 \label{tabref:MC1}
                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
  \hline
\hline
Mean                     & -0.00 & -0.90*   & -0.49 \\
Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
Skew                      &  -0.07 &  -0.16   &  -0.78*\\
Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
Number                 &  1000        &   50        & 50\\

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
This is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citep{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

@@

% The table may be pushed onwards.  Use t (i think).

Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. This apparent stability hides some significant changes that are taking place below the surface.  For example, the Japanese yen records an appreciation of 10.66\% on one day.  Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.  An asterisk is used to denote statistical significance at the level of 95\% confidence.  A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.

@@

The CFTC require traders to categorise themselves as being either commercial (C) or non-commercial (NC).  Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.  These could be seen as natural hedgers.  Non-commercial accounts are generally considered to be speculators.  \citep{FuturesSanders} and \citep{FuturesWang} have examples of previous use of this data in this fashion.

@@
